---
title: "Lab Chapter 2"
output:
  html_document:
    df_print: paged
---
8.
This exercise relates to the College data set, which can be found in
the file College.csv. It contains a number of variables for 777 different
universities and colleges in the US.
Before reading the data into R, it can be viewed in Excel or a text
editor.

(a) Use the read.csv() function to read the data into R. Call the
loaded data college. Make sure that you have the directory set
to the correct location for the data.
```{r}
#Read in file
?read.csv
college <- read.csv(file='College.csv' , header=TRUE )
```

(b) Look at the data using the fix() function. You should notice
that the first column is just the name of each university.We don't
really want R to treat this as data. However, it may be handy to
have these names for later. Try the following commands:
  rownames (college )=college [,1]
  fix (college )
You should see that there is now a row.names column with the
name of each university recorded. This means that R has given
each row a name corresponding to the appropriate university. R
will not try to perform calculations on the row names. However,
we still need to eliminate the first column in the data where the
names are stored. Try
  college =college [,-1]
  fix(college)
Now you should see that the first data column is Private. Note
that another column labeled row.names now appears before the
Private column. However, this is not a data column but rather
the name that R is giving to each row.
```{r}
#View dataset
fix(college) #opens the file in a window, NOTE YOU CAN CHANGE THE VALUES HERE!
rownames(college)=college[,1] #adds a column of rownames to the college matrix, that won't be operated on
fix(college)
college<-college[,-1] #removes the (now excessive) first column
fix(college)
```


(c)
  i. 
Use the summary() function to produce a numerical summary
of the variables in the data set.
  ii.
Use the pairs() function to produce a scatterplot matrix of
the first ten columns or variables of the data. Recall that
you can reference the first ten columns of a matrix A using
A[,1:10].
  iii. 
Use the plot() function to produce side-by-side boxplots of
Outstate versus Private.
  iv. 
Create a new qualitative variable, called Elite, by binning
the Top10perc variable. We are going to divide universities
into two groups based on whether or not the proportion
of students coming from the top 10% of their high school
classes exceeds 50%.
  Elite =rep ("No",nrow(college ))
  Elite [college$Top10perc >50]=" Yes"
  Elite =as.factor (Elite)
  college =data.frame(college ,Elite)
Use the summary() function to see how many elite universities
there are. Now use the plot() function to produce
side-by-side boxplots of Outstate versus Elite.
  v. 
Use the hist() function to produce some histograms with
differing numbers of bins for a few of the quantitative variables.
You may find the command par(mfrow=c(2,2)) useful:
it will divide the print window into four regions so that four
plots can be made simultaneously. Modifying the arguments
to this function will divide the screen in other ways.
  vi. 
Continue exploring the data, and provide a brief summary
of what you discover.
```{r}
summary(college) #gives stat info about each variable
pairs(college[,1:10]) #gives scatter matrix between all indicated variables (columns, thus variables, can be sepcified), as in the following line
plot(college[,1] , college[,9]) #y=outstate=column 9 VS x=private=column 1 ##plot y vs x ###plot automatically gives boxplot when variable is categorical
#Binning
Elite = rep("No" , nrow(college)) #replicates a vector, of all "No"s, nrow times
Elite[college$Top10perc>50]="Yes" #makes the value at which college$Top10perc>50 to be yes
Elite = as.factor(Elite) #This was a vector of No's and Yes's, now split up as a 'factor' with 2 'levels'
#Modify dataset
college = data.frame(college,Elite) #appends the factor Elite to the end of college
#Observe new dataset
summary(college)
plot(college[,19],college[,9]) #y=outstate=column 9 VS x=elite=column 19 ##plot y vs x
hist(college[,2])
par(mfrow=c(4,5)) #splits the view window into 20 parts ##use dev.off() to clear plots, when there isn't any open graphical device
for ( variable in c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18) ){
  hist(college[,variable] , main=paste('Histogram of',variable.names(college)[variable]) , xlab=variable.names(college)[variable])
}
```

9.
This exercise involves the Auto data set studied in the lab. Make sure
that the missing values have been removed from the data.
```{r}
#Read in and view dataset
auto <- read.csv(file='Auto.csv' , header=TRUE , na.strings="?")
View(auto)
#Two ways of taking care of data. This first is averaging the column value. We have enough observations where we can just remove it though. So commenting out the first method:
#auto$horsepower = ifelse(is.na(auto$horsepower),
#                   ave(auto$horsepower, FUN = function(x) mean(x, na.rm = TRUE)),
#                   auto$horsepower)
  #Removing the rows at which there are missing values in the horsepower column. This can be generalized to a whole dataset using the mod function when using is.na(auto) to get the proper row
y <- which(is.na(auto$horsepower))
auto <- auto[-y,]
```

(a) Which of the predictors are quantitative, and which are qualitative?
```{r,echo=FALSE}
print("The last two columns, which are auto[,8], origin, and auto[,9], name, are categorical.")
```

(b) What is the range of each quantitative predictor? You can answer
this using the range() function.
```{r,echo=FALSE}
for ( variable in c(1:7) ){
  y=range(auto[,variable])
  print(paste('The range of' , variable.names(auto)[variable] , 'is' , y[1] , 'to' , y[2] ))
}
```

(c) What is the mean and standard deviation of each quantitative
predictor?
```{r,echo=FALSE}
for ( variable in c(1:7) ){
  y=mean(auto[,variable])
  print(paste(paste(paste('The mean of' , variable.names(auto)[variable]) , 'is') , y) )
}
for ( variable in c(1:7) ){
  y=sd(auto[,variable])
  print(paste(paste(paste('The standard deviation of' , variable.names(auto)[variable]) , 'is') , y) )
}
```

(d) Now remove the 10th through 85th observations. What is the
range, mean, and standard deviation of each predictor in the
subset of the data that remains?
```{r}
#Creating new dataset for (d)
auto_d <- auto[-c(10:85),]
```
```{r,echo=FALSE}
for ( variable in c(1:7) ){
  y=range(auto_d[,variable])
  print(paste('The range of' , variable.names(auto_d)[variable] , 'is' , y[1] , 'to' , y[2] ))
  y=mean(auto_d[,variable])
  print(paste(paste(paste('The mean of' , variable.names(auto_d)[variable]) , 'is') , y) )
  y=sd(auto_d[,variable])
  print(paste(paste(paste('The standard deviation of' , variable.names(auto_d)[variable]) , 'is') , y) )
}
```

(e) Using the full data set, investigate the predictors graphically,
using scatterplots or other tools of your choice. Create some plots
highlighting the relationships among the predictors. Comment
on your findings.
```{r}
pairs(auto)
```
```{r, echo=FALSE}
print('Considering mpg, there seems to be some correlations with: cylinders, displacement, horsepower, and weight, only slightly with year.')
```
```{r}
pairs(auto[,c(1,2,3,4,5,7)])
```
```{r, echo=FALSE}
print('See part f of the question, below.')
print('Displacement, horsepower, and weight seem to have high correlation.')
```
```{r}
pairs(auto[,c(3,4,5)])
cor(auto[,3],auto[,4])
cor(auto[,4],auto[,5])
cor(auto[,3],auto[,5])
```
```{r, echo=FALSE}
print('Cylinders has correlation with displacement, horsepower, weight, and slightly with acceleration.')
```
```{r}
pairs(auto[,c(2,3,4,5,6)])
cor(auto[,2],auto[,3])
cor(auto[,2],auto[,4])
cor(auto[,2],auto[,5])
cor(auto[,2],auto[,6])
```

(f) Suppose that we wish to predict gas mileage (mpg) on the basis
of the other variables. Do your plots suggest that any of the
other variables might be useful in predicting mpg? Justify your
answer.
```{r}
#Plotting other variables vs mpg
par(mfrow=c(2,3))
for( V in c(2,3,4,5,7)){
  plot(auto[,V],auto[,1] , xlab=variable.names(auto)[V] , ylab=variable.names(auto)[1])
}
for( V in c(2,3,4,5,7)){
  print(cor(auto[,V],auto[,1]))
}
```
```{r, echo=FALSE}
print('Since there is a decently high (in absolue value) correlation between mpg and weight, displacement, horsepower, and cylinders, (in that order), these variables may be useful in predicting mpg. Perhaps surprisingly, year has only a slightly positive correlation.')
```

10. This exercise involves the Boston housing data set.

(a) To begin, load in the Boston data set. The Boston data set is
part of the MASS library in R.
  library (MASS)
Now the data set is contained in the object Boston.
  Boston
Read about the data set:
  ?Boston
How many rows are in this data set? How many columns? What
do the rows and columns represent?
```{r}
#As directed
library(MASS)
?Boston
dataset <- Boston
```
```{r, echo=FALSE}
print( paste("The number of rows in the dataset Boston is", nrow(Boston),"and the number of columns is", ncol(Boston),"." ) )
```
```{r}
attach(dataset) #makes it so that you can call columns by their variables, ie. crim rather than dataset$crim or dataset[, {crim index} ]
  # checking that it worked. it did!
  print(dataset[1,1])
  print(crim[1])
```
```{r, echo=FALSE}
print("Rows represent different areas of Boston. Towns or neighborhoods")
```
```{r,echo=FALSE}
for( i in 1:ncol(dataset) ){
  if (i==1) {
    y="per capita crime rate by town."
  } else if (i==2) {
    y="proportion of residential land zoned for lots over 25,000 sq.ft."
  } else if (i==3) {
    y="proportion of non-retail business acres per town."
  } else if (i==4) {
    y="Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)."
  } else if (i==5) {
    y="nitrogen oxides concentration (parts per 10 million)."
  } else if (i==6) {
    y="average number of rooms per dwelling."
  } else if (i==7) {
    y="proportion of owner-occupied units built prior to 1940."
  } else if (i==8) {
    y="weighted mean of distances to five Boston employment centres."
  } else if (i==9) {
    y="index of accessibility to radial highways."
  } else if (i==10) {
    y="full-value property-tax rate per $10,000."
  } else if (i==11){
    y="pupil-teacher ratio by town."
  } else if (i==12){
    y="1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town."
  } else if (i==13){
    y="lower status of the population (percent)."
  } else {
    y="median value of owner-occupied homes in $1000s."
  }
  print( paste("The",i,"th column", colnames(dataset[i]),"represents:",y))
}
```


(b) Make some pairwise scatterplots of the predictors (columns) in
this data set. Describe your findings.
```{r, echo=FALSE}
pairs(dataset)
print("Some observations are:")
print("Potentially exponential results are:")
print("crim vs [age, dis,medv]")
print("zn vs [nox, age, dis, lstat]")
print("indus vs dis")
print("nox vs [age, dis]")
print("rm vs [age,dis]")
print("age vs lstat")
print("dis vs [black,lstat]")
print("black,medv")
print("lstat,medv.")
print("Rest of question skipped because next chapter will have better analytic commands introduced, anyway.")
```

(c) Are any of the predictors associated with per capita crime rate?
If so, explain the relationship.
```{r}
par(mfrow=c(4,4))
for( i in 1:ncol(dataset) ){
  plot(dataset[,i],crim,xlab=colnames(dataset)[i])
}
```
```{r, echo=FALSE}
print("When the region's proportion of residential land zoned for lots is close to zero, crime varies from very low to very high, otherwise it's very low.")
print("When the region's proportion of non-retail business acres per town is at about 17, crime varies from very low to very high, otherwise it's very low. ")
print("In regions without the Charles River, crime varies from very low to very high, otherwise it's very low.")
print("In regions where nitrous oxides conentration is low and high, the crime rate is low. The closer the concentration is to 0.7, the more variation there is in crime rate. There are also a few fairly high crime rate-areas with concentration of 0.6.")
print("In regions where average rooms per dwelling is high and low, crime rate is low; otherwise, crime rate varies.")
print("In regions where there is a high proportion of owner-occupied units built before 1940, the crime rate varies more, especially past 90. Otherwise, it is very low for the most part.")
print("In regions where the distance is very close to Boston employment centers, the crime rate varies from very low to very high. It is very low otherwise.")
print("In regions where the index of accessibility to radial highways is high, the crime rate varies from very low to very high. It is very low, otherwise.")
print("In regions where the property-tax rate per 10,000 is about 675, the crime rate varies from very low to very high, and is very low otherwise.")
print("In regions where the pupil-teacher ratio is just over 20, the crime rate varies from very low to very high. It is very low otherwise.")
print("There is a higher crime rate per capita of lower-proportional black-populated regions than higher-proportional black-populated regions.")
print("There is a small crime rate in regions with less than to percent of the lower status of the population.")
print("There is a very low crime rate in regions where the median value of owner-occupied homes is more than $30,000.")
print("Generally, there seems to be many strong (non-linear) correlations between regions and crime rates. This not surprising to me, the environment is certainly the most influential factor on crime.")
```

(d) Do any of the suburbs of Boston appear to have particularly
high crime rates? Tax rates? Pupil-teacher ratios? Comment on
the range of each predictor.
```{r}
hist(crim) #What's going on when crime rate is greater than 10? than 20?
ten_crim <- crim[crim>10] #looks only at regions with crim>20
twenty_crim  <- crim[crim>20] #looks only at regions with crim>20
par(mfrow=c(1,3))
hist(crim)
hist(ten_crim)
hist(twenty_crim)
```
```{r, echo=FALSE}
print("The vast majority of regions has a very low crime rate, about 450 of the 506 regions. The next highest bin, from 10-10 crime rate, contains 36 regions. There are very few regions with higher rates than a crime rate of 30, and there is no region with a crime rate above 90. So, it would be worth considering just these regions.") #do that, then. if you want?
```
```{r}
hist(tax)
```
```{r, echo=FALSE}
print("The range of tax rates with the most regions is 650-700. There are very few regions greater than 700. However, there are no regions with a tax between 500 and 650.")
```
```{r,echo=FALSE}
hist(ptratio)
print("The ratio with the most (by far) regions is 20.")
for(v in 1:ncol(dataset)){
    print(range(dataset[,v]))
}
```

(e) How many of the suburbs in this data set bound the Charles
river?
```{r,echo=FALSE}
#Since the value is 1 if tract bounds river, I can use sum():
print(paste("The number of regions that bound the Charles river is",sum(chas),"."))
```

(f) What is the median pupil-teacher ratio among the towns in this
data set?
```{r,echo=FALSE}
print(paste("The median of t the pupil-teacher ratio among towns is",median(ptratio),"."))
```

(g) Which suburb of Boston has lowest median value of owneroccupied
homes? What are the values of the other predictors
for that suburb, and how do those values compare to the overall
ranges for those predictors? Comment on your findings.
```{r,echo=FALSE}
lowest_medv = dataset[medv==min(medv),]
for ( l in 1:(ncol(dataset)) ){
  x=range(dataset[,l])
  y=range(lowest_medv[,l])
  print(paste("The range of",colnames(dataset)[l],"is",x[1],"to",x[2],"."))
  print(paste("The range of",colnames(dataset)[l],"at the lowest value of owner-occupied homes is",y[1],"to",y[2],"."))
}
```
```{r, echo=FALSE}
print("Where the value of homes is lowest: crime is high or very high, there is no land zoned for 25,000 sq. ft. lots, the proportion of non-retail businesses is fairly low (based on range, so not so useful), they don't bound the Charles River, the nitrgen oxide concentration is fairly center (based on range, so not that useful), the average number of rooms per dwelling fairly center (based on range, so not that useful), the proportion of owner-occupied units built before 1940 is 100, the distance from Boston employment centers is low, their accessibility to radial highways is extremely high, the property tax is extremely high, the pupil-teacher ratio is as most areas, proportion of blacks is extremely high, and the lower status of population is fairly high.")
```

(h) In this data set, how many of the suburbs average more than
seven rooms per dwelling? More than eight rooms per dwelling?
Comment on the suburbs that average more than eight rooms
per dwelling.
```{r,echo=FALSE}
hist(rm)
sevenplus <- dataset[rm>=7,]
hist(sevenplus[,6],labels=TRUE)
eightplus <- dataset[rm>=8,]
hist(eightplus[,6])
print(paste( "There are",nrow(sevenplus),"suburbs with an average of more than seven rooms per dwelling, and",nrow(eightplus),"suburbs with an average of more than eight rooms per dwelling." ))
```